True versus apparent shapes of bow shocks and bow waves

### 2017-11-05 papers/Paper1/sec-conic.tex
Redo projected quadrics figures with new notation

%The transformation between the conic parameters and the characteristic radii are given by:
%\begin{align}
%R_c &= \frac{b^2}{a} \\
%R_{90} &= b\left[\pm\left(1 - \frac{(a \mp R_0)^2}{a^2}\right)\right]^{1/2}\label{eq:r90-conic}
%\end{align}ut 
%In (\ref{eq:r90-conic}), the positive sign correspond to an elliptic quadric, while the negative sign to an hyperbolic quadric.
%The inverse transformation of (\ref{eq:Rc-conic}) and (\ref{eq:r90-conic}) is given by:
%\begin{align}
%\frac{a}{R_0} &= \frac{A}{2A-B^2} \label{eq:a-conic}\\
%\frac{|b|}{R_0} &= \frac{A}{\left|2A-B^2\right|^{1/2}}\\
%\tan\theta_c &= \frac{2A-B^2}{\left|B^2 - 2A\right|^{1/2}} \label{eq:thc-conic} \\
%\mathrm{where:~} B &\equiv \frac{R_{90}}{R_0} \\
%#A &\equiv \frac{R_c}{R_0}
%\end{align}

@@

We now apply the machinery of \S~\ref{sec:projection} to find the
projected shape of a quadric bow on the plane of the sky.  The
intrinsic 3D shape of the shell is given by rotating
equations~\eqref{eq:par-xy} about the \(x\)-axis, but it is more
convenient to first transform to a reference frame where the origin is
at the center of the conic section:
\begin{equation}
  \label{eq:xyz-XYZ}
  (X, Y, Z) = (x - x_0, y, z) . 
\end{equation}
In this new frame, the quadric shape is
\begin{gather}
  \label{eq:quadric-XYZ}
  \begin{aligned}
    X &= a\Cos(t) \\ 
    Y &= b \Sin(t)\cos\phi \\
    Z &= b \Sin(t)\sin\phi
  \end{aligned}
\end{gather}
The azimuth of the tangent line as a function of inclination and
parametric variable is then found from equations (\ref{eq:alpha},
\ref{eq:tanphi}) to be
\begin{equation}
  \label{eq:phit-quadric}
  \sin\phi_{\T} = \frac{b \Cos(t)} {a \Sin(t)} \tan i \ .
\end{equation}
Combining equations~(\ref{eq:Trans}, \ref{eq:quadric-XYZ},
\ref{eq:phit-quadric}) gives the observer-frame cartesian
plane-of-sky coordinates of the tangent line:
\begin{gather}
  \label{eq:conic-projected-XY}
  \begin{aligned}
    X_{\T}' & = \frac{\Cos(t)}{a\cos i}
    \left(a^2\cos^2 i + \sigma b^2\sin^2 i\right)
    \\
    Y_{\T}' &= b\Sin(t)
    \left(
      1 - \frac{b^2 \Cos^2(t)}{a^2 \Sin^2(t)}
      \tan^2 i\right)^{1/2}
  \end{aligned}
\end{gather}
We wish to show that the projected shape is a conic section of the
same variety (ellipse or hyperbola) as the one that generated the
original quadric.  If this \emph{were} true, then it would be possible
to write the plane-of-sky coordinates as
\begin{gather}
  \begin{aligned}
    X_{\T}' &= a'\Cos( t')  \\
    Y_{\T}' &= b'\Sin (t')  . 
  \end{aligned}\label{eq:conic-projected-XY-conic}
\end{gather}
Comparing equations (\ref{eq:conic-projected-XY}) and
(\ref{eq:conic-projected-XY-conic}), we find after some algebra that
the two forms for \((X_{\T}', Y_{\T}')\) are indeed consistent, with
the equivalences:
%
\newcommand\fQi{\ensuremath{f_{\scriptscriptstyle \Q,i}}}
\begin{align}
  \label{eq:a-prime}
  a' &= a \fQi \cos i  \\
  \label{eq:b-prime}
  b' &= b \\
  t' &= \Cos^{-1} \left[ \fQi \Cos(t) \right]  \ ,
  % t' &= \Cos^{-1} \left[  \frac{a'\Cos(t)}{a\cos i}\right] \ .
\end{align}
where for convenience we define the quadric projection factor:
\begin{equation}
  \label{eq:fQi-factor}
  \fQi = \left( 1 + \Q \tan^2 i \right)^{1/2} \ .
\end{equation}
This demonstrates the original claim that the projected shape is also
a conic section, which means that we can re-use the previous
equations~(\ref{eq:R0-from-Q-a-x0}--\ref{eq:Tq-from-Pi-Lambda}) with
primed quantities substituted for unprimed ones.  From
equations~(\ref{eq:Tq}, \ref{eq:a-prime}, \ref{eq:b-prime}) it follows
that the quadric parameter of the projected shape is
\begin{equation}
  \label{eq:Q-prime}
  \Q' = \frac{ \Q } { \fQi^2 \cos^2 i} \ .
\end{equation}
Finally, we transform the projected reference frame back to be
centered on the star again:
\begin{equation}
  \label{eq:XYZ-xyz-prime}
  (x_{\T}', y_{\T}') = (X_{\T}' + x_0', Y_{\T}') \ , 
\end{equation}
where the projected quadric displacement \(x_0'\) follows from simple
foreshortening:
\begin{equation}
  \label{eq:x0-prime}
  x_0' = x_0 \cos i \ .
\end{equation}

Finally, using $D' = D\cos i$ and substituting $a'$ instead of $a$ in equation (\ref{eq:x0}) we find that:

\begin{align}
R'_0 &= \pm a' + x_0\cos i \\
R'_0 &= \pm\left(a^2\cos^2 i \pm b^2\sin^2 i\right)^{1/2} +  (R_0 \mp a)\cos i
\end{align}

Defining  $f(i;\theta_c)\equiv\left(1\pm\tan^2\theta_c\tan^2i\right)^{1/2}$ we find that:
\begin{align}
\frac{R'_0}{D'}=\frac{R_0}{D}\left(1\pm \tilde{R}_c\cot^2\theta_c(f(i;\theta_c)-1) \right)
\label{eq:qprime}
\end{align}
%Where $A\equiv \frac{R_c}{R_0}$
The Radius of curvature in the observer's frame is given by $R'_c=\frac{b'^2}{a'}$. Then:
\begin{align}
  \tilde{R}_c'\equiv  \frac{R'_c}{R'_0} &= \frac{\tilde{R}_c}{\cos^2 i f(i;\theta_c)\frac{q'}{q}}
  \label{eq:Aprime}\\
  \tan\theta'_c & = \frac{\tan\theta_c}{\cos i f(i;\theta_c)} \\
  \tilde{R}'_{90} &= \left(\frac{2\tilde{R}_cf(i;\theta_c) \mp \tan^2\theta_c\frac{q'}{q}}{\frac{q'}{q}}\right)^{1/2}\frac{\sec i}{f(i;\theta_c)}
\end{align}
where $q=\frac{R_0}{D}$ and $q' = \frac{R'_0}{D'}$.

When $R'_{90}$ is measurable, diagnostic diagrams like figure (\ref{fig:projected-R90-vs-Rc}) can be used to compare with observations.  
\begin{figure*}
  \centering
  \setkeys{Gin}{width=0.48\linewidth}
  \begin{tabular}{@{}ll@{}}
    (a) & (b) \\
    \includegraphics{figs/projected-Rc-vs-i}
        & \includegraphics{figs/projected-R0-vs-i} \\
    (c) & (d) \\
    \includegraphics{figs/projected-R90-vs-i}
    & \includegraphics{figs/projected-R90-vs-Rc}
  \end{tabular}
  \caption[]{Effects of projection on quadrics of revolution:
    variation with inclination of characteristic radii.  Different
    line colors correspond to different quadric shapes (see key),
    while variation in line width corresponds to variation in the
    ``true'' dimensionless radius of curvature, \(\tilde{R}_c\), at
    the apex. Vertical gray rectangles show quartiles of \(|\sin i|\),
    which will be equally populated for an isotropic distribution of
    orientations. (a)~Projected radius of curvature, normalized by
    projected stand-off distance: \( \tilde{R}_c' = R'_c /
    R'_0\). (b)~Projected stand-off distance: \(R_0' / R_0\).
    (c)~Projected perpendicular radius, normalized by projected
    stand-off distance: \( \tilde{R}_{90}' = R'_{90} /
    R'_0\). (d)~Diagnostic diagram for dimensionless projected radii:
    \(\tilde{R}_{90}'\) versus \(\tilde{R}_c'\).  Shading indicates
    different classes of quadrics: hyperboloids (white), prolate
    spheroids (dark gray), and oblate spheroids (light gray), with the
    limiting case of paraboloids shown by the thin black line.}
  \label{fig:quadric-projection}
\end{figure*}

>>>>

%The transformation between the conic parameters and the characteristic radii are given by:
%\begin{align}
%R_c &= \frac{b^2}{a} \\
%R_{90} &= b\left[\pm\left(1 - \frac{(a \mp R_0)^2}{a^2}\right)\right]^{1/2}\label{eq:r90-conic}
%\end{align}ut 
%In (\ref{eq:r90-conic}), the positive sign correspond to an elliptic quadric, while the negative sign to an hyperbolic quadric.
%The inverse transformation of (\ref{eq:Rc-conic}) and (\ref{eq:r90-conic}) is given by:
%\begin{align}
%\frac{a}{R_0} &= \frac{A}{2A-B^2} \label{eq:a-conic}\\
%\frac{|b|}{R_0} &= \frac{A}{\left|2A-B^2\right|^{1/2}}\\
%\tan\theta_c &= \frac{2A-B^2}{\left|B^2 - 2A\right|^{1/2}} \label{eq:thc-conic} \\
%\mathrm{where:~} B &\equiv \frac{R_{90}}{R_0} \\
%#A &\equiv \frac{R_c}{R_0}
%\end{align}
\begin{figure*}
  \centering
  \setkeys{Gin}{width=0.48\linewidth}
  \begin{tabular}{@{}ll@{}}
    (a) & (b) \\
    \includegraphics{figs/projected-Rc-vs-i}
        & \includegraphics{figs/projected-R0-vs-i} \\
    (c) & (d) \\
    \includegraphics{figs/projected-R90-vs-i}
    & \includegraphics{figs/projected-R90-vs-Rc}
  \end{tabular}
  \caption[]{Effects of projection on quadrics of revolution:
    variation with inclination of bow size and shape.  Different line
    colors correspond to varying quadric parameter, \(\Q\), (see key),
    while variation in line width corresponds to variation in the
    ``true'' planitude, \(\Pi\), or apex radius of curvature. Vertical
    gray rectangles show quartiles of \(|\sin i|\), which will be
    equally populated for an isotropic distribution of
    orientations. (a)~Projected planitude: \(\Pi'\). (b)~Projected/true
    star-apex distance: \(R_0' / R_0\).  (c)~Projected alatude:
    \(\Lambda'\). (d)~Diagnostic diagram for
    dimensionless projected radii: \(\tilde{R}_{90}'\) versus
    \(\tilde{R}_c'\).  Shading indicates different classes of
    quadrics: hyperboloids (white), prolate spheroids (dark gray), and
    oblate spheroids (light gray), with the limiting case of
    paraboloids shown by the thin black line.}
  \label{fig:quadric-projection}
\end{figure*}

@@

We now apply the machinery of \S~\ref{sec:projection} to find the
projected shape of a quadric bow on the plane of the sky.  The
intrinsic 3D shape of the shell is given by rotating
equations~\eqref{eq:par-xy} about the \(x\)-axis, but it is more
convenient to first transform to a reference frame where the origin is
at the center of the conic section:
\begin{equation}
  \label{eq:xyz-XYZ}
  (X, Y, Z) = (x - x_0, y, z) . 
\end{equation}
In this new frame, the quadric shape is
\begin{gather}
  \label{eq:quadric-XYZ}
  \begin{aligned}
    X &= a\Cos(t) \\ 
    Y &= b \Sin(t)\cos\phi \\
    Z &= b \Sin(t)\sin\phi
  \end{aligned}
\end{gather}
The azimuth of the tangent line as a function of inclination and
parametric variable is then found from equations (\ref{eq:alpha},
\ref{eq:tanphi}) to be
\begin{equation}
  \label{eq:phit-quadric}
  \sin\phi_{\T} = \frac{b \Cos(t)} {a \Sin(t)} \tan i \ .
\end{equation}
Combining equations~(\ref{eq:Trans}, \ref{eq:quadric-XYZ},
\ref{eq:phit-quadric}) gives the observer-frame cartesian
plane-of-sky coordinates of the tangent line:
\begin{gather}
  \label{eq:conic-projected-XY}
  \begin{aligned}
    X_{\T}' & = \frac{\Cos(t)}{a\cos i}
    \left(a^2\cos^2 i + \sigma b^2\sin^2 i\right)
    \\
    Y_{\T}' &= b\Sin(t)
    \left(
      1 - \frac{b^2 \Cos^2(t)}{a^2 \Sin^2(t)}
      \tan^2 i\right)^{1/2}
  \end{aligned}
\end{gather}
We wish to show that this projected shape is a conic section of the
same variety (ellipse or hyperbola) as the one that generated the
original quadric.  If this \emph{were} true, then it would be possible
to write the plane-of-sky coordinates as
\begin{gather}
  \begin{aligned}
    X_{\T}' &= a'\Cos( t')  \\
    Y_{\T}' &= b'\Sin (t')  . 
  \end{aligned}\label{eq:conic-projected-XY-conic}
\end{gather}
Comparing equations (\ref{eq:conic-projected-XY}) and
(\ref{eq:conic-projected-XY-conic}), we find after some algebra that
the two forms for \((X_{\T}', Y_{\T}')\) are indeed consistent, with
the equivalences:
%
\newcommand\fQi{\ensuremath{f_{\scriptscriptstyle \Q,i}}}
\begin{align}
  \label{eq:a-prime}
  a' &= a \fQi \cos i  \\
  \label{eq:b-prime}
  b' &= b \\
  t' &= \Cos^{-1} \left[ \fQi \Cos(t) \right]  \ ,
  % t' &= \Cos^{-1} \left[  \frac{a'\Cos(t)}{a\cos i}\right] \ .
\end{align}
where for convenience we define the quadric projection factor:
\begin{equation}
  \label{eq:fQi-factor}
  \fQi = \left( 1 + \Q \tan^2 i \right)^{1/2} \ .
\end{equation}
This demonstrates the original claim that the projected shape is also
a conic section, which means that we can re-use the previous
equations~(\ref{eq:R0-from-Q-a-x0}--\ref{eq:Tq-from-Pi-Lambda}) with
primed quantities substituted for unprimed ones.  From
equations~(\ref{eq:Tq}, \ref{eq:a-prime}, \ref{eq:b-prime}) it follows
that the quadric parameter of the projected shape is
\begin{equation}
  \label{eq:Q-prime}
  \Q' = \frac{ \Q } { \fQi^2 \cos^2 i} \ .
\end{equation}
Finally, we transform the projected reference frame back to be
centered on the star again:
\begin{equation}
  \label{eq:XYZ-xyz-prime}
  (x_{\T}', y_{\T}') = (X_{\T}' + x_0', Y_{\T}') \ , 
\end{equation}
where the projected quadric displacement \(x_0'\) follows from simple
foreshortening:
\begin{equation}
  \label{eq:x0-prime}
  x_0' = x_0 \cos i \ .
\end{equation}

The projection of the apex distance then follows from the primed
version of equation~\eqref{eq:R0-from-Q-a-x0} as
\begin{equation}
  \label{eq:R0-prime}
  \frac{R_0'}{R_0} =
  \cos i \left[ 1 + \frac{\Pi} {\Q} \bigl(\fQi - 1\bigr) \right]
\end{equation}
and the projected planitude and alatude can be calculated as
\begin{align}
  \label{eq:Pi-prime}
  \Pi' &= \frac{\Pi}{(R_0'/R_0)\, \fQi \cos i } \\
  \label{eq:Lambda-prime}
  \Lambda' &= \left( 2 \Pi' - \Q' \right)^{1/2} \ .
\end{align}

% Finally, using $D' = D\cos i$ and substituting $a'$ instead of $a$ in equation (\ref{eq:x0}) we find that:

% \begin{align}
% R'_0 &= \pm a' + x_0\cos i \\
% R'_0 &= \pm\left(a^2\cos^2 i \pm b^2\sin^2 i\right)^{1/2} +  (R_0 \mp a)\cos i
% \end{align}

% Defining  $f(i;\theta_c)\equiv\left(1\pm\tan^2\theta_c\tan^2i\right)^{1/2}$ we find that:
% \begin{align}
% \frac{R'_0}{D'}=\frac{R_0}{D}\left(1\pm \tilde{R}_c\cot^2\theta_c(f(i;\theta_c)-1) \right)
% \label{eq:qprime}
% \end{align}
% %Where $A\equiv \frac{R_c}{R_0}$
% The Radius of curvature in the observer's frame is given by $R'_c=\frac{b'^2}{a'}$. Then:
% \begin{align}
%   \tilde{R}_c'\equiv  \frac{R'_c}{R'_0} &= \frac{\tilde{R}_c}{\cos^2 i f(i;\theta_c)\frac{q'}{q}}
%   \label{eq:Aprime}\\
%   \tan\theta'_c & = \frac{\tan\theta_c}{\cos i f(i;\theta_c)} \\
%   \tilde{R}'_{90} &= \left(\frac{2\tilde{R}_cf(i;\theta_c) \mp \tan^2\theta_c\frac{q'}{q}}{\frac{q'}{q}}\right)^{1/2}\frac{\sec i}{f(i;\theta_c)}
% \end{align}
% where $q=\frac{R_0}{D}$ and $q' = \frac{R'_0}{D'}$.

% When $R'_{90}$ is measurable, diagnostic diagrams like figure (\ref{fig:projected-R90-vs-Rc}) can be used to compare with observations.
